Neighborhoods and Felony Disenfranchisement: The Case of New York City

### 2019-05-20 markdown/kmorris_thesis.Rmd
Show that ballot undercount is not associated with lost voters

These individuals are identified by matching Department of Corrections and Community Supervisions data to the voter file using the methodology developed by Burch [-@Burch2010; -@Burch2011]. I use first, middle, and last names, and dates of birth, to join these two datasets.^[Matching on first and last names and dates of birth can result in false positive matches, especially in large states like New York. Following @Meredith2013, I report estimates of the potential impact of these false positives in the Appendix.] Figure \ref{fig:citymap} shows where these lost voters lived before going to prison, with city council districts also included. There were `r comma(readRDS("./temp/lost_count_nyc.rds"))` such lost voters within New York City as of the 2017 general election, and `r comma(nrow(readRDS("./temp/ids_of_lost_voters.rds")))` statewide.

@@

After matching neighborhoods, I use a simple regression to test whether turnout in the 2017 mayoral election was different in areas with lost voters. Census tract and block group level turnout rates are calculated using the geocoded voter file. Each voter's record indicates whether the voter participated in the 2017 general election,^[The New York State voter file is not without errors. For instance, the voter file erroneously does not record the author as participating in the 2017 primary election. However, insofar as these errors are not geographically correlated, there is little reason to believe they impact the results of this analysis. Unfortunately, testing the distribution of these errors is not possible.] which are then aggregated to estimate the number of ballots cast in each neighborhood. The number of ballots cast is divided the voting age population.

@@

\newpage
# References {-}

>>>>

These individuals are identified by matching Department of Corrections and Community Supervisions data to the voter file using the methodology developed by Burch [-@Burch2010; -@Burch2011]. I use first, middle, and last names, and dates of birth, to join these two datasets.^[Matching on first and last names and dates of birth can result in false positive matches, especially in large states like New York. Following @Meredith2013, I report estimates of the potential impact of these false positives in Appendix XX.] Figure \ref{fig:citymap} shows where these lost voters lived before going to prison, with city council districts also included. There were `r comma(readRDS("./temp/lost_count_nyc.rds"))` such lost voters within New York City as of the 2017 general election, and `r comma(nrow(readRDS("./temp/ids_of_lost_voters.rds")))` statewide.

@@

After matching neighborhoods, I use a simple regression to test whether turnout in the 2017 mayoral election was different in areas with lost voters. Census tract and block group level turnout rates are calculated using the geocoded voter file. Each voter's record indicates whether the voter participated in the 2017 general election, which are then aggregated to estimate the number of ballots cast in each neighborhood.^[The New York registered voter file does not align exactly with results reported by the city. The voter file indicates that just XX voters cast a ballot in the 2017 mayoral election, but the Board of Elections reports that XX votes were cast. In [Appendix A], I demonstrate that there is no relationship between lost voters and underreporting of cast ballots at the precinct level, which means these reporting anamolies are unlikely to impact this analysis.] The number of ballots cast is divided the voting age population.

@@

\newpage
# References {-}
<div id="refs"></div>

\newpage
# Appendix A {-}

As noted above, the turnout figures reported in the registered voter file do not align with the results reported by the New York City Board of Elections. According to the official results, XX ballots were cast in the 2017 mayoral election. The registered voter file reports that just XX voters cast a ballot. This is not necessarily evidence of poor management by the Board of Elections. Some voters choose not to make their voter registration information publicly available (such as domestic violence survivors and officers of the court). Where more ballots are recorded by the BOE than in the voter file, turnout calculated from the voter file will be artificially lower. If the voter file undercount rate is systematically worse in neighborhoods with lost voters, this would pose a serious challenge to the validity of the results reported in the body of this paper.

To test whether there is a relationship between lost voters and voter file undercount, I examine the undercount rate at the precinct level. The voter file indicates the home precinct of each voter, and the Board of Elections publishes election results at the precinct level. Because the census does not report statistics at the precinct level, I use block group information instead. Where precincts cross block group lines, the precinct is assigned the characteristics of all block groups in which it is located weighted by the number of voters from each block group. The undercount rate is calculated as the number of votes derived from the voter file divided by the number reported in BOE results.

```{r undercount-reg-chunk, echo = F, results = "asis", message = F}
load("./temp/compare_ballots_off_results.RData")

stargazer(ed_reg,
          header = F,
          type = "latex",
          se = list(ed_comp_ses),
          covariate.labels = c("Lost Voters", "Median Income (Thousands of Dollars)", "Percent Latino", "Percent Non-Hispanic Black", "Percent Non-Hispanic White",
                               "Percent With Some College", "Median Age", "Registration Rate", "Percent Democrats", "Percent Noncitizen", "Percent Won by City Council Representative"),
          dep.var.labels = c("Undercount Rate"),
          title = "(\\#tab:uc-reg) Registered Voter File Ballot Undercount",
          table.placement = "H",
          style = "ajps",
          omit.stat = c("f", "ser"),
          notes = "Robust standard errors (clustered by assembly district) in parentheses.",
          omit = "boro")
```

As Table \ref{tab:uc-reg} makes clear, there is very little relationship (p = 0.47) between the number of lost voters in a precinct and the ballot undercount rate. Although researchers should be somewhat wary using turnout rates derived from New York State's registered voter file, there is no evidence that the reporting error impact this analysis.
